The Timing and Likelihood of the Origin of Life from Earth's Impact History

### 2023-04-25 ms.tex
word smithing appendix

Recently, \citet{Citron_2022} performed smoothed-particle hydrodynamics impact simulations over a wide range of impact angles, masses and velocities to estimate the impact properties that vaporize the ocean. Figure \ref{fig:energy_for_ocean_vap}a shows their simulation results for the change in the atmosphere's and ocean's internal energy ($\Delta IE_\mathrm{atmo}$) as a function of impact kinetic energy ($E_\mathrm{imp}$), along with log-linear extrapolations for each impact angle. Figure \ref{fig:energy_for_ocean_vap} gives the same information, but the y-axis is instead the fraction of the impactor's energy that heats the atmosphere and ocean. For the most probable impact angle of $45^\circ$ about $\sim 10\%$ of the impactors kinetic energy heats the atmosphere ($f_\mathrm{E,vap} = 0.1$), so we nominally adopt this value in the main text to determine which collisions vaporize the ocean.

Energy delivery to the atmosphere/ocean appears to depend on impact angle, mass and velocity (Figure \ref{fig:energy_for_ocean_vap}). Additionally, all of the \citet{Citron_2022} simulations are far more massive than the minimum threshold for ocean-vaporization, so we must rely on extrapolations. Therefore, our assumption of a constant $f_\mathrm{E,vap} = 0.1$ is an over-simplification. To remedy this shortcoming, we considered building a parameterization for $f_\mathrm{E,vap}$ that depends on multiple impact properties (e.g., impact angle) using the \citet{Citron_2022} simulation results, but this was unsuccessful because the calculations do not consider a wide enough parameter space.

@@

The evaluate the sensitivity of our results to an assumed constant $f_\mathrm{E,vap} = 0.1$, we recomputed the likelihood and timing of a life-starting impact using $f_\mathrm{E,vap} = 0.025$ and $f_\mathrm{E,vap} = 0.25$ (Figure \ref{fig:probabilities_of_impacts_sens} and \ref{fig:timing_sensitivity}). We chose these values because they are possible lower and upper bounds based on the \citet{Citron_2022} simulations (Figure \ref{fig:energy_for_ocean_vap}b). The probability of a impact that produces substantial prebiotic molecules without later ocean vaporization decreases with increasing $f_\mathrm{E,vap}$ (Figure \ref{fig:probabilities_of_impacts_sens}). For example, for $f_\mathrm{E,vap} = 0.1$, the probability of a $>4 \times 10^{20}$ kg impact (\citet{Wogan_2023} optimistic case) without subsequent ocean vaporization is 92\%. For $f_\mathrm{E,vap} = 0.025$ and $f_\mathrm{E,vap} = 0.25$ the probabilities are instead 99.9\% and 60\%, respectively.

>>>>

Recently, \citet{Citron_2022} performed smoothed-particle hydrodynamics impact simulations over a range of impact angles, masses and velocities to estimate the impact properties that can vaporize an ocean. Figure \ref{fig:energy_for_ocean_vap}a shows their simulation results for the change in the atmosphere's and ocean's internal energy ($\Delta \mathrm{IE}_\mathrm{atmo}$) as a function of impact kinetic energy ($\mathrm{E}_\mathrm{imp}$), along with log-linear extrapolations for each impact angle. Figure \ref{fig:energy_for_ocean_vap} gives the same information, but the y-axis is instead the fraction of the impactor's energy that heats the atmosphere and ocean. For the most probable impact angle of $45^\circ$ about $\sim 10\%$ of the impactors kinetic energy heats the atmosphere ($f_\mathrm{E,vap} = 0.1$), so we nominally adopt this value in the main text to determine which collisions vaporize the ocean.

Energy delivery to the atmosphere/ocean appears to depend on impact angle, mass and velocity (Figure \ref{fig:energy_for_ocean_vap}). Additionally, all of the \citet{Citron_2022} simulations are far more massive than the minimum threshold for ocean-vaporization, so we must rely on extrapolations. Therefore, our assumption of a constant $f_\mathrm{E,vap} = 0.1$ is an over-simplification. To remedy this shortcoming, we considered building a parameterization for $f_\mathrm{E,vap}$ that depends on multiple impact properties (e.g., impact angle) using the \citet{Citron_2022} simulation results, but this was unsuccessful because their calculations do not consider a wide enough parameter space.

@@

The evaluate the sensitivity of our results to an assumed constant $f_\mathrm{E,vap} = 0.1$, we recomputed the likelihood and timing of a life-starting impact using $f_\mathrm{E,vap} = 0.025$ and $f_\mathrm{E,vap} = 0.25$ (Figure \ref{fig:probabilities_of_impacts_sens} and \ref{fig:timing_sensitivity}). We chose these values because they are reasonable lower and upper bounds based on the \citet{Citron_2022} simulations (Figure \ref{fig:energy_for_ocean_vap}b). The probability of a impact that produces substantial prebiotic molecules without later ocean vaporization decreases with increasing $f_\mathrm{E,vap}$ (Figure \ref{fig:probabilities_of_impacts_sens}). For example, for $f_\mathrm{E,vap} = 0.1$, the probability of a $>4 \times 10^{20}$ kg impact (\citet{Wogan_2023} optimistic case) without subsequent ocean vaporization is 92\%. For $f_\mathrm{E,vap} = 0.025$ and $f_\mathrm{E,vap} = 0.25$ the probabilities are instead 99.9\% and 60\%, respectively.
